---
title: "Litter Fall Analysis"
author: "Kevin Styers"
date: "July 6, 2019"
output:
  prettydoc::html_pretty:
    toc: true
    theme: architect
    highlight: github
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning=FALSE,error = TRUE)
library(ggplot2)
theme_set(theme_light())
```

```{r initial clean, include=FALSE, echo=FALSE}
# setwd("X:/1_Databases")

library(plotly)

# Stack Data
#library(neonUtilities)
#neonUtilities::stackByTable("NEON_litterfall.zip")

# Grab Data
library(data.table)
litter <- data.table::fread("X:/1_Databases/NEON_litterfall/stackedFiles/ltr_massdata.csv")
litter <- litter %>% mutate(year = substring(weighDate,1,nchar(weighDate) -6))
litter1 <- litter
# Select Data
litter <- litter %>%
    select(-remarks) %>%
    na.omit()


# Convert columns into usable data types
library(tidyverse)
library(magrittr)
cols <- c("domainID", "siteID", "plotID", "trapID","functionalGroup","dryMass","qaDryMass","measuredBy")
litter <- litter %>% mutate_at(cols, funs(factor(.)))
litter$weighDate <- as.Date(litter$weighDate)
litter$setDate <- as.Date(litter$setDate)
litter$collectDate <- as.Date(litter$collectDate)
litter$ovenStartDate <- as.Date(litter$ovenStartDate)
litter$ovenEndDate <- as.Date(litter$ovenEndDate)
litter$dryMass <- as.numeric(as.character(litter$dryMass, na.rm = TRUE))
litter<- litter %>% mutate(dayOfYear = yday(weighDate))
litter <- litter %>%
    mutate(totalSampleTime = collectDate - setDate)

# Mutate data

# # Total Sample Time
litter <- litter %>%
    mutate(totalSampleTime = collectDate - setDate)
# # Sum of dryWeight for each functional Group
litter <- select(litter, domainID, siteID, plotID, trapID,year,dayOfYear, weighDate, setDate, collectDate, ovenStartDate, ovenEndDate, fieldSampleID, massSampleID, functionalGroup, dryMass, qaDryMass, measuredBy,totalSampleTime,measuredBy)

groups <- c("Leaves", "Twigs/branches","Other", "Woody material","Seeds","Needles","Mixed","Flowers")
```

# Introduction  

The NEON Project collects leaf litter in 1m x 1m 'litter traps,' spread across a site's forested plots. Technicians periodically collect, sort, dry, and weigh the litter trap's contents by functional group. I've done some exploratory analysis to look at some macro patterns without a specific goal. I use tidy techniques and `ggplot2` with `plotly` to transform and visualize the NEON Data.


# NEON's Total Litter Weights 

## by Functional Group  

This first plot shows the NEON Project's total litter collection weights by functional group and Domain ID. The Y-Axis shows the total weigh in grams, the X-Axis filters the data by fuctional group. if you hover your cursor you can see precise weights for functional groups with that Domain.

```{r plot1, echo=FALSE, fig.width= 8}
selectionVariable <- litter %>%
    select(year,domainID,functionalGroup,dryMass) %>%
    filter(year > "2015") %>%
    group_by(domainID,functionalGroup) %>%
    summarize(siteFuncWeight = sum(dryMass))

# Test a plot 
plot <- ggplot(selectionVariable, aes(x=functionalGroup, y=siteFuncWeight, fill = domainID)) +
    geom_col() +
    labs(title = "",
         subtitle = "",
         fill='DomainID:',
         y = "Mass (g)",
         x = "Functional Group") +
    theme(axis.text.x = element_text(angle = 270)) +
    scale_x_discrete(name ="Functional Group", 
                    limits=c("Flowers","Other","Woody material","Seeds","Needles","Mixed","Twigs/branches","Leaves"))

ggplotly(plot)   
```

## by DomainID  

Now, I've switched the x-axis and color variables to show the total of weight each functional group by Domain

```{r plot2, echo=FALSE, fig.width= 8}
d5 <- litter %>%
    filter(domainID == "D05") 

selectionVariable <- litter %>%
    filter(year > "2015" & year < "2019") %>%
    select(domainID,functionalGroup,dryMass) %>%
    group_by(domainID,functionalGroup) %>%
    summarize(siteFuncWeight = sum(dryMass))
# groups <- c("Leaves", "Twigs/branches","Other", "Woody material","Seeds","Needles","Mixed","Flowers")

plot <- ggplot(selectionVariable, aes(x=domainID, y=siteFuncWeight, fill = functionalGroup)) +
    geom_col() +    
    labs(title = "",
         fill='DomainID:',
         y = "Mass (g)",
         x = "") +
    theme(axis.text.x = element_text(angle = 270),
          axis.title.y = element_text(angle = 0))

ggplotly(plot) 
```

## by Total Sample Time in Days  

```{r plot4, echo=F,fig.width= 8}
litter1 <- litter %>%
    select(domainID,trapID,setDate,collectDate,totalSampleTime) %>%
    distinct() %>% 
    group_by(domainID,trapID,setDate) %>%
    summarize(TOTALset = sum(totalSampleTime))%>%
    group_by(domainID)%>%
    summarize(totalSet = sum(TOTALset))
litter1$totalSet <- as.integer(as.character(litter1$totalSet))
plot <- ggplot(litter1, aes(x=domainID, y=totalSet, fill=domainID))+
        geom_col() +    
    labs(title = "",
         fill='DomainID:',
         y = "Total Sample Days",
         x = "") +
    theme(axis.text.x = element_text(angle = 270),
          axis.title.y = element_text(angle = 0))
ggplotly(plot)
```


## by Date

The X-Axis is now the Day of Year. The Y-Axis is the total weight (grams) for that Day of Year.  
```{r plot3, echo=F, fig.width= 8,warning=FALSE}
# group Weight and Date
weighDay <- litter %>%
    filter(year > "2015" & year < "2019") %>%
    select(year,dayOfYear,functionalGroup,dryMass)%>%
    group_by(year, dayOfYear,functionalGroup) %>%
    summarize(dayMass = sum(dryMass))
# weighDay$weighDate <- as.Date(weighDay$weighDate)

weighDay$dayOfYear <- as.numeric(as.character(weighDay$dayOfYear))

plot <- ggplot(weighDay,aes(x=dayOfYear,y=dayMass,color=functionalGroup))+
    geom_line() +
    labs(title = "NEON's Total Litter Weights by Functional Group",
       fill='DomainID:',
       y = "",
       x = "Day of Year",
       color = "Functional Group") +
    facet_wrap(~year,ncol=1)
ggplotly(plot)
```


## by Technician

This plot shows technicians with highest sampled litter weight in grams.   

```{r plot5, echo=F, fig.width= 8,warning=FALSE}
tfts <- litter %>%
  filter(year > "2015" & year <"2019")%>%
  select(dryMass, measuredBy) %>%
  group_by(measuredBy)%>%
  summarize(totalMass = sum(dryMass))

df <-tfts[order(-tfts$totalMass),]
tftsHead <- head(df, n=10)

plot <- ggplot(tftsHead, aes(x=measuredBy, y=totalMass, fill=measuredBy))+
        geom_col() +    
    labs(title = "",
         fill='Measured By:',
         y = "Total Mass Sampled(g)",
         x = "") +
    theme(axis.text.x = element_text(angle = 270),
          axis.title.y = element_text(angle = 0))

ggplotly(plot)

```
